Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-16 paper/ms.tex
bona fide

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.95\textwidth]{f8.pdf}
		\vspace{-0.1cm}
		\caption{
			{\bf Properties of CQVs identified by our search}.
      CQVs are mostly pre-main-sequence M dwarfs, younger than
      $\approx$150 Myr, with rotation periods faster than
      $\approx$1\,day.  The \ngoods\ bonafide CQVs in
      Table~\ref{tab:thetable} are the dark blue circles; \nmaybes\
      ambiguous CQV candidates are light blue circles.  Unresolved
      binaries are flagged either if the Gaia DR3 radial velocity
      scatter exceeded 20\,\kms, or if Gaia ${\rm RUWE}_{\rm DR3}$$>$2
      and multiple photometric signals were present in the TESS light
      curve.  The top panels show the \nstarssearched\ target stars
      with 120-second cadence TESS data as the shaded gray background;
      darker regions correspond to a larger relative number of
      searched stars.  The lower-left panel compares the
      rotation--color distribution of CQVs against the rotation
      periods of K and M dwarfs in the Pleiades from
      \citet{2016AJ....152..114R}.  The lower-middle panel plots the
      derived corotation radii $R_{\rm cr} = (GM/\Omega^2)^{1/3}$ in
      units of stellar radii against the measured CQV periods, in
      units of hours.  Ages in the final panel are known from cluster
			membership.
		}
		\label{fig:catalogscatter}
	\end{center}
\end{figure*}

@@

In our target sample, we found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bonafide CQVs are listed in Table~\ref{tab:thetable}.  This table also
includes an additional \nmaybes\ ambiguous CQVs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but
two of the \nallcands\ objects based on association memberships; we
also derived temperatures and radii using SED fitting, and then
inferred stellar masses by interpolating against stellar evolutionary
models.  We caution that our selection function was not
volume-limited: the TESS 120-second stellar sample had a heterogeneous
selection function which may have been biased in favor of young stars
over field stars.  Previous work however has shown that $\approx$1-3\%
of M dwarfs younger than $\approx$100\,Myr show the CQV phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G}.

@@

\begin{enumerate}[leftmargin=*]
	%
    \item CQVs are quasiperiodic.  The mean periods remain fixed over
      the $>$1{,}000-cycle baseline of available observations; but the
      light curve shapes always evolve (Figure~\ref{fig:evoln}).
  %
    \item The same CQV can show multiple periods simultaneously.  LP
      12-502, for instance, showed dips with four distinct periods
      within $\pm 0.3\%$ of its fundamental period, sometimes
      simultaneously, and each lasting for up to 50 cycles
      (Figure~\ref{fig:lpriver0}).
  %
    \item CQVs evolve over timescales that are both secular
      ($>$100\,cycles) and impulsive ($<$1 cycle).  Dip growth seems
      to happen over durations of at least ten cycles, and slow dip
      decay can also occur.  ``State-switches'' correspond to dips to
      collapsing instantaneously, and are almost always linked with
      observed optical flares.  This suggests that the occulting
      material is sensitive to sudden changes in the magnetic field.
  %
    \item The rate of dip evolution can be used to place a
      model-dependent constraint on how much material is either being
      accreted or ejected during the state changes
      (Section~\ref{subsec:massflux}).  Order of magnitude estimates
      require at least an asteroid belt's worth of dust
      ($10^{-4}$\,$M_\oplus$) over $10^8$ years, or
      $\approx$$10^{-2}$\,$M_\oplus$ if the occulting material is gas.
  %
    \item The CQV phenomenon persists for $\gtrsim$150\,Myr, based on
      the existence of multiple CQVs in AB Dor, the Pleiades, and
      Psc-Eri (Section~\ref{subsec:starprops}).  It may even extend to
      200\,Myr, based on the one CQV we found in the Carina Near
      moving group (TIC~294328887; $\approx$200\,Myr).  However the
      lack of detected CQVs in the Hyades and Praesepe suggests that
      the lifetime of the phenomenon is limited to the first few
      hundred million years.
	%
    \item The duty cycle for CQVs seems to be $\approx$$75$\%, based
      on the fraction of bonafide CQVs that either turned on or turned
      off during TESS re-observations, two years after the initial
      observation.
  %
    \item Most CQVs are M dwarfs with masses
      0.1-0.5\,$M_\odot$.  Two sources, TIC~405754448 and
      TIC~405910546, have masses that appear to exceed the M dwarf
      limit.  However both are potentially binaries, and this may
      confuse our ability to accurate identify the source of the CQV
      signal (Section~\ref{subsec:massive}).  We encourage additional
      scrutiny of these objects in future work.
  %
    \item Surprising analogs to the CQVs exist in magnetic B stars,
      but only those with multipolar field topologies.  Since most
      magnetic B stars have dipolar magnetic fields, this suggests
      that the CQV dips and warps are similarly being sculpted by the
      stellar magnetic fields, and that the magnetic fields themselves
      are probably also multipolar.
  %
    \item The closest CQVs to the Sun are at distances of 15-20\,pc,
      and the brightest have $V$$\approx$12 ($J$$\approx$7.5).  We
      have found most of them in this work, since our CQV sample was
      $\gtrsim$80\% complete within 30\,pc.  The lack of CQVs in the
      volume-complete $<$15\,pc sample of 0.1-0.3\,$M_\odot$ stars
      analyzed by \citet{2021AJ....161...63W} is consistent with this
      estimate.  Expanding our analysis of the TESS data to the full
      frame images would yield a truly volume-limited selection
      function, and would expand the CQV census by about a factor of
      two within 50\,pc, and by a factor of ten within 100\,pc.
\end{enumerate}

@@

\startlongtable
\begin{deluxetable}{rrrrrrlrllrrrrrr}
\tabletypesize{\footnotesize}
%\tabletypesize{\scriptsize}
\tablecaption{Bonafide, candidate, and debunked complex quasiperiodic
  variables from the TESS 2-minute data. {\bf For internal
  review, full versions are available here
  \url{https://www.dropbox.com/scl/fo/twn4s9ckbevf75jqhtoy1/h?rlkey=t5cn8cx2uoc2ptdm9e570kp1b&dl=0}} \label{tab:thetable}}
%\toprule
%TIC & $T$ & $d$ & BP-RP & RUWE & $P$ & assoc & age & teff & rstar & mstar & rcr & qual & Nsectors \\
%\midrule
%\endhead
\startdata
  TIC & $T$ & $d$ & $\bprp$ & RUWE & $P$ & Assoc & Age & $T_{\rm eff}$ & $R_\star$ & $M_\star$ & $R$$_{\rm cr}$ & $P_{\rm sec}$ & Quality & Bin & $N_{\rm sector}$ \\
  -- &   mag & pc &    mag &      -- &        hr &   -- &        Myr &  K                  & $R_\odot$ & $M_\odot$ & $R_\star$         & hr                     & -- & -- & -- \\
\hline
\input{20230613_LGB_RJ_CPV_TABLE_selfnapplied_rounded_data}
\enddata
\tablecomments{This table includes \ngoods\ CQVs (\texttt{Quality}
  flag 1), \nmaybes\ ambiguous CQVs (\texttt{Quality} flag 0), and
  \ndebunked\ impostors (\texttt{Quality} flag -1).  The three-bit
  binarity flag ``Bin'' is for Gaia DR3
  \texttt{radial\_velocity\_error} outliers (bit 1), Gaia DR3
  \texttt{ruwe} outliers (bit 2), and stars with multiple TESS periods
  (bit 3).  The machine-readable version, available online, includes
  additional columns for the Gaia DR2 and DR3 source identifiers, as
  well as the stellar parameter uncertainties.  The age uncertainties
  are typically $\approx \pm 10\%$, but can be asymmetric.  {\bf The
  median temperature, radius, and mass uncertainties are $\pm XXX$\,K,
  $\pm XX$\%, and $\pm YY$\% respectively}.  $N_{\rm sector}$ denotes
  the number of TESS sectors for which {\it any} data are expected to
  be acquired between July~2018 and Oct~2024.  This number is
  generally greater than the number of sectors for which 120-second
  cadence data exist.   Association names and provenance follow
  conventions adopted by \citet{2018ApJ...856...23G}: 
	ABDMG: AB~Doradus moving group \citep{2015MNRAS.454..593B}.
	ARG: Argus \citep{2019ApJ...870...27Z}.
	%APER: $\alpha$~Persei open cluster \citep{2023AJ....166...14B}.
	BPMG: $\beta$~Pic moving group \citep{2015MNRAS.454..593B}.
  CARN: Carina Near moving group \citep{2006ApJ...649L.115Z}.
	COL: Columba \citep{2015MNRAS.454..593B}.
	EPSC: $\epsilon$ Chamaeleontis \citep{2013MNRAS.435.1325M}.
	LCC: Lower Centaurus Crux \citep{2016MNRAS.461..794P}.
	PERI: Pisces-Eridani \citep{2019AJ....158...77C}.
	PLE: Pleiades \citep{2015ApJ...813..108D}.
	TAU: Taurus \citep{1995ApJS..101..117K}.
	THA: Tucana-Horologium assocation \citep{2015MNRAS.454..593B}.
	TWA: TW Hydrae assocation \citep{2015MNRAS.454..593B}.
	UCL: Upper Centaurus Lupus \citep{2016MNRAS.461..794P}.
	USCO: Upper Scorpius \citep{2016MNRAS.461..794P}.
	The ``(?)'' string denotes low-confidence membership.
}
\end{deluxetable}
\clearpage

>>>>

\begin{figure*}[!t]
	\begin{center}
		\centering
		\includegraphics[width=0.95\textwidth]{f8.pdf}
		\vspace{-0.1cm}
		\caption{
			{\bf Properties of CQVs identified by our search}.
      CQVs are mostly pre-main-sequence M dwarfs, younger than
      $\approx$150 Myr, with rotation periods faster than
      $\approx$1\,day.  The \ngoods\ bona fide CQVs in
      Table~\ref{tab:thetable} are the dark blue circles; \nmaybes\
      ambiguous CQV candidates are light blue circles.  Unresolved
      binaries are flagged either if the Gaia DR3 radial velocity
      scatter exceeded 20\,\kms, or if Gaia ${\rm RUWE}_{\rm DR3}$$>$2
      and multiple photometric signals were present in the TESS light
      curve.  The top panels show the \nstarssearched\ target stars
      with 120-second cadence TESS data as the shaded gray background;
      darker regions correspond to a larger relative number of
      searched stars.  The lower-left panel compares the
      rotation--color distribution of CQVs against the rotation
      periods of K and M dwarfs in the Pleiades from
      \citet{2016AJ....152..114R}.  The lower-middle panel plots the
      derived corotation radii $R_{\rm cr} = (GM/\Omega^2)^{1/3}$ in
      units of stellar radii against the measured CQV periods, in
      units of hours.  Ages in the final panel are known from cluster
			membership.
		}
		\label{fig:catalogscatter}
	\end{center}
\end{figure*}

@@

In our target sample, we found \ngoods\ objects that showed complex
quasiperiodic behavior over at least one TESS sector.  These \ngoods\
bona fide CQVs are listed in Table~\ref{tab:thetable}.  This table also
includes an additional \nmaybes\ ambiguous CQVs, whose designation is
less certain, and \ndebunked\ impostors.  We inferred ages for all but
two of the \nallcands\ objects based on association memberships; we
also derived temperatures and radii using SED fitting, and then
inferred stellar masses by interpolating against stellar evolutionary
models.  We caution that our selection function was not
volume-limited: the TESS 120-second stellar sample had a heterogeneous
selection function which may have been biased in favor of young stars
over field stars.  Previous work however has shown that $\approx$1-3\%
of M dwarfs younger than $\approx$100\,Myr show the CQV phenomenon
\citep{2016AJ....152..114R,2022AJ....163..144G}.

@@

\begin{enumerate}[leftmargin=*]
	%
    \item CQVs are quasiperiodic.  The mean periods remain fixed over
      the $>$1{,}000-cycle baseline of available observations; but the
      light curve shapes always evolve (Figure~\ref{fig:evoln}).
  %
    \item The same CQV can show multiple periods simultaneously.  LP
      12-502, for instance, showed dips with four distinct periods
      within $\pm 0.3\%$ of its fundamental period, sometimes
      simultaneously, and each lasting for up to 50 cycles
      (Figure~\ref{fig:lpriver0}).
  %
    \item CQVs evolve over timescales that are both secular
      ($>$100\,cycles) and impulsive ($<$1 cycle).  Dip growth seems
      to happen over durations of at least ten cycles, and slow dip
      decay can also occur.  ``State-switches'' correspond to dips to
      collapsing instantaneously, and are almost always linked with
      observed optical flares.  This suggests that the occulting
      material is sensitive to sudden changes in the magnetic field.
  %
    \item The rate of dip evolution can be used to place a
      model-dependent constraint on how much material is either being
      accreted or ejected during the state changes
      (Section~\ref{subsec:massflux}).  Order of magnitude estimates
      require at least an asteroid belt's worth of dust
      ($10^{-4}$\,$M_\oplus$) over $10^8$ years, or
      $\approx$$10^{-2}$\,$M_\oplus$ if the occulting material is gas.
  %
    \item The CQV phenomenon persists for $\gtrsim$150\,Myr, based on
      the existence of multiple CQVs in AB Dor, the Pleiades, and
      Psc-Eri (Section~\ref{subsec:starprops}).  It may even extend to
      200\,Myr, based on the one CQV we found in the Carina Near
      moving group (TIC~294328887; $\approx$200\,Myr).  However the
      lack of detected CQVs in the Hyades and Praesepe suggests that
      the lifetime of the phenomenon is limited to the first few
      hundred million years.
	%
    \item The duty cycle for CQVs seems to be $\approx$$75$\%, based
      on the fraction of bona fide CQVs that either turned on or turned
      off during TESS re-observations, two years after the initial
      observation.
  %
    \item Most CQVs are M dwarfs with masses
      0.1-0.5\,$M_\odot$.  Two sources, TIC~405754448 and
      TIC~405910546, have masses that appear to exceed the M dwarf
      limit.  However both are potentially binaries, and this may
      confuse our ability to accurate identify the source of the CQV
      signal (Section~\ref{subsec:massive}).  We encourage additional
      scrutiny of these objects in future work.
  %
    \item Surprising analogs to the CQVs exist in magnetic B stars,
      but only those with multipolar field topologies.  Since most
      magnetic B stars have dipolar magnetic fields, this suggests
      that the CQV dips and warps are similarly being sculpted by the
      stellar magnetic fields, and that the magnetic fields themselves
      are probably also multipolar.
  %
    \item The closest CQVs to the Sun are at distances of 15-20\,pc,
      and the brightest have $V$$\approx$12 ($J$$\approx$7.5).  We
      have found most of them in this work, since our CQV sample was
      $\gtrsim$80\% complete within 30\,pc.  The lack of CQVs in the
      volume-complete $<$15\,pc sample of 0.1-0.3\,$M_\odot$ stars
      analyzed by \citet{2021AJ....161...63W} is consistent with this
      estimate.  Expanding our analysis of the TESS data to the full
      frame images would yield a truly volume-limited selection
      function, and would expand the CQV census by about a factor of
      two within 50\,pc, and by a factor of ten within 100\,pc.
\end{enumerate}

@@

\startlongtable
\begin{deluxetable}{rrrrrrlrllrrrrrr}
\tabletypesize{\footnotesize}
%\tabletypesize{\scriptsize}
\tablecaption{Bona fide, candidate, and debunked complex quasiperiodic
  variables from the TESS 2-minute data. {\bf For internal
  review, full versions are available here
  \url{https://www.dropbox.com/scl/fo/twn4s9ckbevf75jqhtoy1/h?rlkey=t5cn8cx2uoc2ptdm9e570kp1b&dl=0}} \label{tab:thetable}}
%\toprule
%TIC & $T$ & $d$ & BP-RP & RUWE & $P$ & assoc & age & teff & rstar & mstar & rcr & qual & Nsectors \\
%\midrule
%\endhead
\startdata
  TIC & $T$ & $d$ & $\bprp$ & RUWE & $P$ & Assoc & Age & $T_{\rm eff}$ & $R_\star$ & $M_\star$ & $R$$_{\rm cr}$ & $P_{\rm sec}$ & Quality & Bin & $N_{\rm sector}$ \\
  -- &   mag & pc &    mag &      -- &        hr &   -- &        Myr &  K                  & $R_\odot$ & $M_\odot$ & $R_\star$         & hr                     & -- & -- & -- \\
\hline
\input{20230613_LGB_RJ_CPV_TABLE_selfnapplied_rounded_data}
\enddata
\tablecomments{This table includes \ngoods\ CQVs (\texttt{Quality}
  flag 1), \nmaybes\ ambiguous CQVs (\texttt{Quality} flag 0), and
  \ndebunked\ impostors (\texttt{Quality} flag -1).  The three-bit
  binarity flag ``Bin'' is for Gaia DR3
  \texttt{radial\_velocity\_error} outliers (bit 1), Gaia DR3
  \texttt{ruwe} outliers (bit 2), and stars with multiple TESS periods
  (bit 3).  The machine-readable version, available online, includes
  additional columns for the Gaia DR2 and DR3 source identifiers, as
  well as the stellar parameter uncertainties.  The age uncertainties
  are typically $\approx \pm 10\%$, but can be asymmetric.  {\bf The
  median temperature, radius, and mass uncertainties are $\pm XXX$\,K,
  $\pm XX$\%, and $\pm YY$\% respectively}.  $N_{\rm sector}$ denotes
  the number of TESS sectors for which {\it any} data are expected to
  be acquired between July~2018 and Oct~2024.  This number is
  generally greater than the number of sectors for which 120-second
  cadence data exist.   Association names and provenance follow
  conventions adopted by \citet{2018ApJ...856...23G}: 
	ABDMG: AB~Doradus moving group \citep{2015MNRAS.454..593B}.
	ARG: Argus \citep{2019ApJ...870...27Z}.
	%APER: $\alpha$~Persei open cluster \citep{2023AJ....166...14B}.
	BPMG: $\beta$~Pic moving group \citep{2015MNRAS.454..593B}.
  CARN: Carina Near moving group \citep{2006ApJ...649L.115Z}.
	COL: Columba \citep{2015MNRAS.454..593B}.
	EPSC: $\epsilon$ Chamaeleontis \citep{2013MNRAS.435.1325M}.
	LCC: Lower Centaurus Crux \citep{2016MNRAS.461..794P}.
	PERI: Pisces-Eridani \citep{2019AJ....158...77C}.
	PLE: Pleiades \citep{2015ApJ...813..108D}.
	TAU: Taurus \citep{1995ApJS..101..117K}.
	THA: Tucana-Horologium assocation \citep{2015MNRAS.454..593B}.
	TWA: TW Hydrae assocation \citep{2015MNRAS.454..593B}.
	UCL: Upper Centaurus Lupus \citep{2016MNRAS.461..794P}.
	USCO: Upper Scorpius \citep{2016MNRAS.461..794P}.
	The ``(?)'' string denotes low-confidence membership.
}
\end{deluxetable}
\clearpage
